---
title: "p8105_hw6_ms6826"
output: github_document
date: "2024-11-25"
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

## Problem 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 
```{r}
boot_results=
  weather_df |> 
  modelr::bootstrap(5000) |> 
  mutate(
    strap=map(strap, as_tibble),
    models=map(strap,\(df) lm(tmax~tmin, data=df)),
    r_squared = map_dbl(models, \(model) summary(model)$r.squared), 
    beta0 = map_dbl(models, \(model) coef(model)[1]), 
    beta1 = map_dbl(models, \(model) coef(model)[2]), 
    log_beta_product = log(beta0 * beta1) 
  ) |> 
  select(.id, r_squared, log_beta_product)

boot_results
```

Plot the distribution of your estimates, and describe these in words.
```{r}
boot_results |> 
  ggplot(aes(x=r_squared))+
  geom_histogram(binwidth = 0.001, fill = "skyblue", color = "white") +
  labs(title = "Bootstrap R^2 Distribution", x = "R^2", y = "Count") +
  theme_minimal()

boot_results |> 
  ggplot(aes(x=log_beta_product))+
  geom_histogram(binwidth = 0.01, fill = "red", color = "white") +
  labs(title = "Bootstrap log(β0 * β1) Distribution", x = "log(β0 * β1)", y = "Count") +
  theme_minimal()

```
---The distribution of r^2 appears to be relatively normal, with its center around ~0.913. The distribution of log(β0 * β1) also appears to be normal, with its center closer to 2.01. 

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval 
```{r}
boot_results |> 
  summarize(
    rsq_ci_ll=quantile(r_squared, 0.025),
    rsq_ci_ul=quantile(r_squared, 0.975),
    logp_ci_ll=quantile(log_beta_product, 0.025),
    logp_ci_ul=quantile(log_beta_product, 0.975),
  ) 
```
---The 95% confidence interval for r^2 is (0.894, 0.927) and the 95% confidence interval for log(β0 * β1) is (1.96, 2.06).

## Problem 2

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicide_df=
  read_csv("data/homicide-data.csv") 
  
hom_df=
homicide_df |> 
  mutate(
    city_state=paste(city, state, sep=", "),
    solved = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Closed by arrest" ~ 1,
      disposition == "Open/No arrest" ~ 0),
    victim_age=as.numeric(victim_age)
    ) |> 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
    ) 

```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
balt_reg=
hom_df |> 
  filter(city_state=="Baltimore, MD") |> 
  glm(solved~ victim_age + victim_race + victim_sex, data = _, family=binomial()) |> 
  broom::tidy()

balt_reg_notidy=
hom_df |> 
  filter(city_state=="Baltimore, MD") |> 
  glm(solved~ victim_age + victim_race + victim_sex, data = _, family=binomial())

conf_int_exp=confint.default(balt_reg_notidy) |> exp() 

conf_int_exp[4, ]

balt_reg |> 
  filter(term=="victim_sexMale") |> 
  mutate(
    odds_ratio = exp(estimate)
  ) |> 
  pull(odds_ratio)

```

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
results_df=
  hom_df |> 
  group_by(city_state) |> 
  nest() |> 
  mutate(
    model = map(data, \(x) glm(solved ~ victim_age + victim_race + victim_sex, data=x, family = binomial())),
    tidy_results=map(model, broom::tidy),
    confint_results = map(model, ~ exp(confint.default(.x)))
    ) |> 
 unnest(tidy_results) |>  
  filter(term == "victim_sexMale")  |>   
  mutate(
    odds_ratio = exp(estimate),  
    conf_low = confint_results[[1]][4, 1], 
    conf_high = confint_results[[1]][4, 2]  
  ) |> 
  select(city_state, odds_ratio, conf_low, conf_high) 

print(results_df)
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
MxS CANNOT GET THIS TO ORGANIZE BY OR
```{r}
results_df |>  
  arrange(odds_ratio) |> 
  mutate(city_state = factor(city_state, levels = city_state)) |> 
  ggplot(aes(x=odds_ratio, y=city_state)) +
  geom_point() +
  geom_errorbar(aes(xmin = conf_low, xmax = conf_high), width = 0.2) +
    labs(
    title = "Odds Ratio for Solving Homicides Comparing Male Victims to Female Victims",
    x = "Odds Ratio (with 95% CI)",
    y = "City"
  )
```



## Problem 3
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
bwt_df=
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex=
      case_match(babysex, 
                 1 ~ "male",
                 2 ~ "female"),
    babysex=fct_infreq(babysex),
    frace=
      case_match(frace,
                 1 ~ "white",
                 2 ~ "black",
                 3 ~ "asian",
                 4 ~ "puerto rican",
                 8 ~ "other",
                 9 ~ "unknown"
                  ),
    frace=fct_infreq(frace),
    mrace=
      case_match(mrace,
                 1 ~ "white",
                 2 ~ "black",
                 3 ~ "asian",
                 4 ~ "puerto rican",
                 8 ~ "other"
                  ),
    mrace=fct_infreq(mrace),
    malform=
      case_match(malform,
                 0 ~ "absent",
                 1~ "present"),
    malform=fct_infreq(malform)
  ) 
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Look at data 
```{r}
bwt_df |> 
  ggplot(aes(x=gaweeks, y=bwt))+
  geom_point(alpha=.5)

bwt_df |> 
  ggplot(aes(x=momage, y=bwt))+
  geom_point(alpha=.5)

bwt_df |> 
  ggplot(aes(x=smoken, y=bwt))+
  geom_point(alpha=.5)

```
Fit some models 
```{r}
linear_mod_momage=lm(bwt~momage, data=bwt_df)
linear_mod_smoken=lm(bwt~smoken, data=bwt_df)
linear_mod_gaweeks=lm(bwt~gaweeks, data=bwt_df)

pwl_mod_momage=lm(bwt~momage, data=bwt_df)
pwl_mod_smoken=lm(bwt~smoken, data=bwt_df)
pwl_mod_gaweeks=lm(bwt~gaweeks, data=bwt_df)

smooth_mod_momage=gam(bwt~s(momage), data=bwt_df)
smooth_mod_smoken=gam(bwt~s(smoken), data=bwt_df)
smooth_mod_gaweeks=gam(bwt~ s(gaweeks), data=bwt_df)
```

Look at models
```{r}
bwt_df |> 
  add_predictions(linear_mod_gaweeks) |>
  ggplot(aes(x=gaweeks, y=bwt))+
  geom_point(alpha=.5) +
  geom_line(aes(y=pred), color="red")

bwt_df |> 
  add_predictions(pwl_mod_gaweeks) |>
  ggplot(aes(x=gaweeks, y=bwt))+
  geom_point(alpha=.5) +
  geom_line(aes(y=pred), color="red")


bwt_df |> 
  add_predictions(smooth_mod_gaweeks) |>
  ggplot(aes(x=gaweeks, y=bwt))+
  geom_point(alpha=.5) +
  geom_line(aes(y=pred), color="red")


bwt_df |> 
  add_predictions(linear_mod_momage) |> 
  ggplot(aes(x=momage, y=bwt))+
  geom_point(alpha=.5) +
  geom_line(aes(y=pred), color="red")

bwt_df |> 
  add_predictions(pwl_mod_momage) |> 
  ggplot(aes(x=momage, y=bwt))+
  geom_point(alpha=.5) +
  geom_line(aes(y=pred), color="red")


bwt_df |> 
  add_predictions(smooth_mod_momage) |> 
  ggplot(aes(x=momage, y=bwt))+
  geom_point(alpha=.5) +
  geom_line(aes(y=pred), color="red")


bwt_df |> 
  add_predictions(linear_mod_smoken) |> 
  ggplot(aes(x=smoken, y=bwt))+
  geom_point(alpha=.5) +
   geom_line(aes(y=pred), color="red")
bwt_df |> 
  add_predictions(pwl_mod_smoken) |> 
  ggplot(aes(x=smoken, y=bwt))+
  geom_point(alpha=.5) +
   geom_line(aes(y=pred), color="red")
bwt_df |> 
  add_predictions(smooth_mod_smoken) |> 
  ggplot(aes(x=smoken, y=bwt))+
  geom_point(alpha=.5) +
   geom_line(aes(y=pred), color="red")

```

Fit a model
```{r}
fit=
  gam(bwt~s(momage) + s(gaweeks) + s(smoken), data = bwt_df)

bwt_df |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  ggplot(aes(x=pred, y=resid)) +
  geom_point()

```
---I chose to include the predictors 1)mothers age at delivery, 2)gestational age, and 3)average number of cigarettes smoked per day during pregnancy, because research has shown that these factors affect the health and the birthweight of newborns. I chose a smooth model because I observed that the relationships between birthweight and both mother's age at delivery and average number of cigarettes smoked are not linear, and a smooth model will help account for this non-linearity. This plot of resid vs pred for the smooth model is more promising than the linear model, because the datapoints are less clustered and more spread around the 0 line on the y axis.  

Compare your model to two others:
One using length at birth and gestational age as predictors (main effects only)
```{r}
two_fit=gam(bwt ~ s(blength) + s(gaweeks), data =bwt_df) 

bwt_df |> 
  modelr::add_residuals(two_fit) |> 
  modelr::add_predictions(two_fit) |> 
  ggplot(aes(x=pred, y=resid)) +
  geom_point()

```
--Compared to the model I fit (that includes mothers age at delivery, gestational age, and average number of cigarettes smoked per day during pregnancy), both models are demonstrating some clustering around the 0 y-axis line towards the right of the graphs. Both models also have several points with very large residuals. The model I fit, however, is less clustered than this model. My model also appears to be less hetereoscedastic compared to this model.

One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
int_fit= gam(
  bwt ~ s(blength) + s(bhead) +
    te(blength, bhead) +
    te(blength, by= babysex) +
    te(bhead, by= babysex) +
    te(blength, bhead, by=babysex),
  data = bwt_df
)

bwt_df |> 
  modelr::add_residuals(int_fit) |> 
  modelr::add_predictions(int_fit) |> 
  ggplot(aes(x=pred, y=resid)) +
  geom_point()
```
--Compared to the model I fit, this model's points are more centered around the 0 y-axis line (indicating a better fit). My model also has more clustered data and there are more datapoints with larger residuals (indicating a worse fit). This model, however, appears to be more hetereoscedastic compared to my model.

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r}
cv_df=
  crossv_mc(bwt_df, 100) |> 
  mutate(
    train=map(train, as_tibble),
    test=map(test, as_tibble)
  )

cv_res_df=
  cv_df |> 
  mutate(
    my_mod=map(train, \(x)  gam(bwt~s(momage) + s(gaweeks) + s(smoken), data = x)),
    two_mod=map(train, \(x) gam(bwt ~ s(blength) + s(gaweeks), data =x) ),
    int_mod=map(train, \(x) gam(
  bwt ~ s(blength) + s(bhead) +
    te(blength, bhead) +
    te(blength, by= babysex) +
    te(bhead, by= babysex) +
    te(blength, bhead, by=babysex),
  data = x
))
  ) |> 
  mutate(
    rmse_mine=map2_dbl(my_mod, test, rmse),
    rmse_twoterms=map2_dbl(two_mod, test, rmse),
    rmse_interactions=map2_dbl(int_mod, test, rmse)
  )

cv_res_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x=model, y=rmse))+
  geom_violin()
```
---Based off this graph, since lower RSME indicates a better fit, the model with head circumference, length, sex, and all interactions appears to be the best model. My model has the highest RSME. There is also some  variability in RSME for for model with the interactions compared to the model with just the two terms (length at birth and gestational age).

