---
title: "p8105_hw6_ms6826"
output: github_document
date: "2024-11-25"
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

## Problem 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 
```{r}
boot_results=
  weather_df |> 
  modelr::bootstrap(5000) |> 
  mutate(
    strap=map(strap, as_tibble),
    models=map(strap,\(df) lm(tmax~tmin, data=df)),
    r_squared = map_dbl(models, \(model) summary(model)$r.squared), 
    beta0 = map_dbl(models, \(model) coef(model)[1]), 
    beta1 = map_dbl(models, \(model) coef(model)[2]), 
    log_beta_product = log(beta0 * beta1) 
  ) |> 
  select(.id, r_squared, log_beta_product)

boot_results
```

Plot the distribution of your estimates, and describe these in words.
```{r}
boot_results |> 
  ggplot(aes(x=r_squared))+
  geom_histogram(binwidth = 0.001, fill = "skyblue", color = "white") +
  labs(title = "Bootstrap R^2 Distribution", x = "R^2", y = "Count") +
  theme_minimal()

boot_results |> 
  ggplot(aes(x=log_beta_product))+
  geom_histogram(binwidth = 0.01, fill = "red", color = "white") +
  labs(title = "Bootstrap log(β0 * β1) Distribution", x = "log(β0 * β1)", y = "Count") +
  theme_minimal()

```
---The distribution of r^2 appears to be relatively normal, with its center around ~0.913. The distribution of log(β0 * β1) also appears to be normal, with its center closer to 2.01. 

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval 
```{r}
boot_results |> 
  summarize(
    rsq_ci_ll=quantile(r_squared, 0.025),
    rsq_ci_ul=quantile(r_squared, 0.975),
    logp_ci_ll=quantile(log_beta_product, 0.025),
    logp_ci_ul=quantile(log_beta_product, 0.975),
  ) 
```
---The 95% confidence interval for r^2 is (0.894, 0.927) and the 95% confidence interval for log(β0 * β1) is (1.96, 2.06).

## Problem 2

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicide_df=
  read_csv("data/homicide-data.csv") 
  
hom_df=
homicide_df |> 
  mutate(
    city_state=paste(city, state, sep=", "),
    solved = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Closed by arrest" ~ 1,
      disposition == "Open/No arrest" ~ 0),
    victim_age=as.numeric(victim_age)
    ) |> 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
    ) 

```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
balt_reg=
hom_df |> 
  filter(city_state=="Baltimore, MD") |> 
  glm(solved~ victim_age + victim_race + victim_sex, data = _, family=binomial()) |> 
  broom::tidy()

balt_reg_notidy=
hom_df |> 
  filter(city_state=="Baltimore, MD") |> 
  glm(solved~ victim_age + victim_race + victim_sex, data = _, family=binomial())

conf_int_exp=confint.default(balt_reg_notidy) |> exp() 

conf_int_exp[4, ]

balt_reg |> 
  filter(term=="victim_sexMale") |> 
  mutate(
    odds_ratio = exp(estimate)
  ) |> 
  pull(odds_ratio)

```

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
results_df=
  hom_df |> 
  group_by(city_state) |> 
  nest() |> 
  mutate(
    model = map(data, \(x) glm(solved ~ victim_age + victim_race + victim_sex, data=x, family = binomial())),
    tidy_results=map(model, broom::tidy),
    confint_results = map(model, ~ exp(confint.default(.x)))
    ) |> 
 unnest(tidy_results) |>  
  filter(term == "victim_sexMale")  |>   
  mutate(
    odds_ratio = exp(estimate),  
    conf_low = confint_results[[1]][4, 1], 
    conf_high = confint_results[[1]][4, 2]  
  ) |> 
  select(city_state, odds_ratio, conf_low, conf_high) 

print(results_df)
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r}
results_df |>  
  arrange(odds_ratio) |> 
  mutate(city_state = factor(city_state, levels = city_state)) |> 
  ggplot(aes(x=city_state, y=odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  coord_flip()+
    labs(
    title = "Odds Ratio for Solving Homicides Comparing Male Victims to Female Victims",
    x = "City",
    y = "Odds Ratio (with 95% CI)"
  )
```


## Problem 3
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
bwt_df=
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex=
      case_match(babysex, 
                 1 ~ "male",
                 2 ~ "female"),
    babysex=fct_infreq(babysex),
    frace=
      case_match(frace,
                 1 ~ "white",
                 2 ~ "black",
                 3 ~ "asian",
                 4 ~ "puerto rican",
                 8 ~ "other",
                 9 ~ "unknown"
                  ),
    frace=fct_infreq(frace),
    mrace=
      case_match(mrace,
                 1 ~ "white",
                 2 ~ "black",
                 3 ~ "asian",
                 4 ~ "puerto rican",
                 8 ~ "other"
                  ),
    mrace=fct_infreq(mrace),
    malform=
      case_match(malform,
                 0 ~ "absent",
                 1~ "present"),
    malform=fct_infreq(malform)
  ) 
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
fit=
  lm(bwt~momage + gaweeks + smoken, data = bwt_df)

bwt_df |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  ggplot(aes(x=pred, y=resid)) +
  geom_point()

```
---I chose to include the predictors 1)mothers age at delivery, 2)gestational age, and 3)average number of cigarettes smoked per day during pregnancy, because research has shown that these factors affect the health (and sometimes the birthweight) of newborns. 

Compare your model to two others:
One using length at birth and gestational age as predictors (main effects only)
--COMPARE residuals/predictions GRAPHS
```{r}
fit |> 
  broom::tidy()

bwt_df |> 
  lm(bwt ~ blength + gaweeks, data =_) |> 
  broom::tidy()
```
--COMPARISON: 

One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
fit |> 
  broom::tidy()

bwt_df |> 
  lm(bwt ~ blength + bhead+ babysex + blength*bhead + blength*babysex + bhead*babysex + blength*babysex*bhead , data =_) |> 
  broom::tidy()
```
--COMPARISON: 

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r}
cv_df=
  crossv_mc(bwt_df, 100) |> 
  mutate(
    train=map(train, as_tibble),
    test=map(test, as_tibble)
  )
cv_res_df=
  cv_df |> 
  mutate(
    linear_my_mod=map(train, \(x) lm(bwt~momage + gaweeks + smoken, data=x)),
    linear_two_mod=map(train, \(x) lm(bwt ~ blength + gaweeks, data=x)),
    linear_int_mod=map(train, \(x) lm(bwt ~ blength + bhead+ babysex + blength*bhead + blength*babysex + bhead*babysex + blength*babysex*bhead , data=x))
  ) |> 
  mutate(
    rmse_mine=map2_dbl(linear_my_mod, test, rmse),
    rmse_twoterms=map2_dbl(linear_two_mod, test, rmse),
    rmse_interactions=map2_dbl(linear_int_mod, test, rmse)
  )

cv_res_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x=model, y=rmse))+
  geom_violin()
```


