---
title: "p8105_hw6_ms6826"
output: github_document
date: "2024-11-25"
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

## Problem 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 
```{r}
boot_results=
  weather_df |> 
  modelr::bootstrap(5000) |> 
  mutate(
    strap=map(strap, as_tibble),
    models=map(strap,\(df) lm(tmax~tmin, data=df)),
    r_squared = map_dbl(models, \(model) summary(model)$r.squared), 
    beta0 = map_dbl(models, \(model) coef(model)[1]), 
    beta1 = map_dbl(models, \(model) coef(model)[2]), 
    log_beta_product = log(beta0 * beta1) 
  ) |> 
  select(.id, r_squared, log_beta_product)

boot_results
```

Plot the distribution of your estimates, and describe these in words.
```{r}
boot_results |> 
  ggplot(aes(x=r_squared))+
  geom_histogram(binwidth = 0.001, fill = "skyblue", color = "white") +
  labs(title = "Bootstrap R^2 Distribution", x = "R^2", y = "Count") +
  theme_minimal()

boot_results |> 
  ggplot(aes(x=log_beta_product))+
  geom_histogram(binwidth = 0.01, fill = "red", color = "white") +
  labs(title = "Bootstrap log(β0 * β1) Distribution", x = "log(β0 * β1)", y = "Count") +
  theme_minimal()

```
---The distribution of r^2 appears to be relatively normal, with its center around ~0.913. The distribution of log(β0 * β1) also appears to be normal, with its center closer to 2.01. 

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval 
```{r}
boot_results |> 
  summarize(
    rsq_ci_ll=quantile(r_squared, 0.025),
    rsq_ci_ul=quantile(r_squared, 0.975),
    logp_ci_ll=quantile(log_beta_product, 0.025),
    logp_ci_ul=quantile(log_beta_product, 0.975),
  ) 
```
---The 95% confidence interval for r^2 is (0.894, 0.927) and the 95% confidence interval for log(β0 * β1) is (1.96, 2.06).
